Analyzing the Impact of Social Attributes on Commit Integration Success

### 2017-03-29 Paper/MSRChallenge2017.tex
moved the balance call after a complaint in the build process
changed a link and full name of avg
CLG edits, looking nice

\section{Introduction}
For managers of an open source project, it is important to ensure that the
project consists of high quality contributions.  As social coding sites attract
more users and projects become more popular, they receive ever more
community contributions.  As the number of project contributions increase,
the project must dedicate more time to vetting these contributions. 
%
Over the years, the open source
community has developed various techniques to vet project contributions, such as pull
requests that must be approved by a core team member~\cite{gousios14}. 
This can become a bottleneck, and thus it is beneficial to prioritize certain
contributions, ideally by predicted quality.

Ideally, the decision to include code in
a project is meritocratic, such that only the code is a factor
in the decision~\cite{Scacchi07}.
While one would think that only the complexity of the commit is important, past
studies have shown that the developer making the commit is also an important
indicator for contribution quality~\cite{tsay14icse, tsay14fse}.
Thus, both technical factors, which describe the commit's complexity, and
social factors, which describe the developer's experience, are important when
investigating contribution quality.

Previous work has focused on the predicted acceptance of open source \emph{pull
requests}~\cite{tsay14icse, tsay14fse} and \emph{continuous integration} builds
in a single project.
By contrast,  we are interested in analyzing contribution quality 
in \emph{continuous integration} across multiple open source projects.  While 
pull request
quality is determined through manual inspection, continuous integration quality
is automatically determined through test cases, and thus may lead to different
quality predictors. 
We investigate the factors related to continuous integration build success
across open source projects.
In this study, we analyze a large set of technical and social attributes 
and how these can be used to predict a contribution's quality, measured by the
success of build integration jobs.
We performed this analysis on a large corpus of open source projects
which incorporated continuous integration in their development practice. 
When a contribution to these projects passes integration tests, 
it suggests that the project developers
might have confidence in the changes. 
Thus, we use the commit's build integration status as a proxy for its quality.

@@

We present the approach taken to guide this study by
following the Goal-Question-Metric (GQM) methodology~\cite{Basili84}:

@@

\vspace{1ex}
\noindent \emph{Do social attributes
improve the predictions of a contribution's build success when compared to
predictions from only technical attributes?}

@@

\vspace{1ex}
\noindent\textbf{Metric.} We measured software contribution quality by the
build status of the
commit. We used a decision tree classifier~\cite{Quinlan86} to predict build 
success. Decision trees 
use binary decisions on the
attributes to predict a final class. They are well
established and interpretable. We used a Weka J48 decision tree for readability.
We first built a
decision tree classifier using only technical factors to 
predict build integration
success.  We then included the social factors, to 
identify the degree to which social factors improve the model's predictive
power.
We evaluated the models with 10-fold cross-validation.

@@

The first five factors represent the changes caused by the commit.  The next two
factors cover the size of the project.  These factors represent the commit's complexity
, and the complexity of the project incorporating the commit.

\vspace{1ex}
\noindent\textbf{Social Factors.} We created a web scraper to extract social data
from developer's profiles.
We used the commit hash and the project name in the 
MSR Challenge~\cite{msr17challenge} data set to locate the
commit web page, leveraging \textit{github.com}'s URL standard for projects and
commit hashes.
%  using the syntax
% below. 
% \newline \newline
% \textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
% \newline \newline
%
We scraped the commit author's
\textit{username} from each commit page to
collect the following from the user's public profile:
%(\textit{https://github.com/\textbf{UserName}}). 
%

@@

\vspace{1ex}
\noindent\textbf{Comparing two data sets.} We compiled the gathered data into 
two data sets: one with only the technical
factors and the build status, and another with both technical and social
factors and the build status. We performed this split to analyze how well 
technical factors alone,
and then both technical and social
factors predict build integration status.

@@

74.43\% of the builds in the reduced data set were successful, and thus a naive
classifier that always guesses ``successful'' would succeed 74.43\% of the time.
To penalize misclassification for
both classes equally, we created a second \emph{balanced} data set by randomly 
sub-sampling the successful builds until the two build status were
equal, through a process known as downsampling. We downsampled the entire
data set and then trained decision tree classifiers on both.  We validated the
classifiers with cross-validation. Since the cross-validation is stratified, the
proportion of 50-50 is maintained through the training and testing folds for the
downsampled data. We present results for both
data sets in Section~\ref{sec:successClassification},  and only the full data 
set in Sections~\ref{sec:core} and \ref{sec:followerSuccess}.

@@

\textbf{Kappa statistic}& 0.21 &  0.42 & 0.28  & 0.52 \\
\textbf{Avg Precision}& 0.75 & 0.80  & 0.65  & 0.76 \\
\textbf{Avg Recall}& 0.77 & 0.81 & 0.64 & 0.76 \\
\textbf{Avg F-Measure}& 0.72 & 0.79  & 0.63  & 0.76 \\

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm.  Without downsampling, the 
decision tree shows an increase of 4.09\% \textit{correctly classified} instances when adding social data. On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, which outperformed the baseline of
guessing all builds passed (50.00\%).
After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data.  These results show that
both technical and social data can be used to improve predictions of build 
integration success.

The precision, recall, and
F-measure are calculated over the successful and unsuccessful classes and then
averaged (standard in Weka).  All these measures, along with the Kappa
statistic, increase with social features, which provide evidence that social 
factors improve build success predictions. The mean 
absolute error also decreases, further substantiating the importance 
of social factors
in predicting build success.

@@

The decision tree results suggested that core team members are
important to classifying build success, and thus we investigated the question
directly. 
We classified the build success rate for
core team members and non-core team members using the data set without 
downsampling.  We found that core team members
triggered 89.39\% (331827/371208) of the builds in the data set and 75.58\%
(250808/331827) of those builds
passed.  Non-core team members triggered 10.61\% (39381/371208) of the builds 
in the data set
and 64.70\% (25479/39381) of the builds passed.  The p-value from a Fisher exact
test 0.0000, which means that this difference is extremely unlikely to occur by
chance. These results show core team 
members are more
likely to have a successful build integration than non-core team members. We
hypothesize this occurs because core team members are more familiar with the 
code, and therefore more likely to have their builds pass.

@@

We also investigated whether developers with followers were more likely to have a
successful build integration, since followers may be seen as ``vouching'' for
the developer's abilities.  We therefore investigated the relationship between
follower count and build success rate (on the data without 
downsampling).  We first investigated whether developers with one or more followers
were more likely to pass the build than those without any followers by
removing commits without any author profile
information (20374 builds, reducing the total build count to
350834). Only 0.87\% (3059/350834) of the
contributions in the data set were performed by developers with zero 
followers.  %, leaving 99.13\% (347775/350834) of the contributions 
%performed by developers with at least one follower.
Developers with 0 followers had a successful build commit 68.49\% (2095/3059)
of the time.  Developers with one or more followers were successful 74.78\%
(260056/347775) of the time.  The p-value of this difference, calculated with a
Fisher exact test, is 0.000.  This shows commit success difference between
developers with no followers and developers with at least one follower is
very unlikely to occur by chance.
%
By contrast, developers with a large number of followers (1000 or more),
performed 6.94\% (24347/350834) of the contributions in the 
data set; their
contributions had a 78.61\% (19138/24347) success rate. The p-value calculated
by comparing this success rate to the success rate of developers with no
followers is 0.0000.  The p-value means that that the difference between
developers with 1000 followers and developers with no followers is very unlikely
to occur by chance.

@@

%\paragraph{Internal validity}
In the study, we use Weka, a well known machine-learning toolkit,
to analyze our data.
Like all tools, Weka may contain bugs. 
Nonetheless, it is a community tested open source tool, which mitigates threats to 
internal validity. We also release our scripts, data, and decision tree 
results in the study\footnote{https://goo.gl/CqTX44}.

@@

The most similar related work to this paper is a case study investigating how
social and technical factors affect build
success on a project with 40 developers~\cite{Hassan06}.
In contrast to this work, we focus on a wide range of open source projects and
measure experience through developer profile information.
Other work has investigated the effects of various factors, such as team
organization, on predicting build success in a single
project~\cite{Kwan11,Wolf09}.
Previous investigations have established relationships between both social and 
technical 
attributes and 
different measures of success, such as pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse} or evaluation latency~\cite{Yu15}.
Tsay et al.~\cite{tsay14icse} 
show the importance of the social connection between the developer initiating a
pull request and the project integrators. We look at a broader aspect of 
social factors, and focus on predicting the Travis CI build success.
Gousios et al.~\cite{gousios14} found that the commit change location is
an important indicator to commit acceptance rate, and that including
test cases increases a
commit's acceptance rate~\cite{gousios15}.
Bettenburgh et 
al.~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, while Vasilescu et al.~\cite{vasilescu14} explore how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated into the automatic build process. Yu et al.~\cite{yu16} 
analyze the impact of CI usage in GitHub and its correlation with software 
quality. By contrast with their work, we focus on the relationship between social 
attributes and continuous integration success.

\section{Conclusion}

@@

% use section* for acknowledgement
\section*{Acknowledgments}
This material is based upon work supported by the National Science Foundation
Graduate Research Fellowship under Grant No. DGE-1252522.
The authors gratefully acknowledge the partial support
of the AFRL (AFRL Contract No. FA8750-15-2-0075). Any 
findings or recommendations are those of the author(s) and do not
necessarily reflect the views of the sponsoring agencies.

@@

\balance
\bibliographystyle{plain}
\bibliography{MSRChallenge2017}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references

>>>>

\section{Introduction}
For managers of an open source project, it is important to ensure that the
project consists of high quality contributions.  As social coding sites attract
more users and projects become more popular, they receive ever more
community contributions.  As the number of project contributions increases,
the project must dedicate more time to vetting these contributions. 
%
The open source
community has developed various techniques to vet project contributions, such as
requiring that pull requests be approved by a core team member~\cite{gousios14}. 
This can become a bottleneck, and thus it is beneficial to prioritize certain
contributions, ideally by predicted quality.

Ideally, the decision to include code in
a project is meritocratic, such that only code quality is a factor
in the decision~\cite{Scacchi07}.
However, beyond commit complexity or correctness, past
studies have shown that the social factors describing the committing developer also importantly
predict contribution success~\cite{tsay14icse, tsay14fse}.
Thus, both technical factors, which describe the commit's complexity, and
social factors, which describe the developer's experience, are important when
investigating contribution success (and thus by extension predicted quality).

This previous work has focused on the predicted acceptance of open source \emph{pull
requests}~\cite{tsay14icse, tsay14fse} and \emph{continuous integration} builds
in a single project. %TODO: should there be a ref here?
By contrast,  we are interested in analyzing contribution quality 
in \emph{continuous integration} across multiple open source projects.  While 
pull request
quality is ultimately determined through manual inspection, continuous
integration uses test cases to automatically assess commit quality,
and thus may provide different quality predictors. 
We investigate the factors related to continuous integration build success
across open source projects.
In this study, we analyze a large set of technical and social attributes 
and how these can be used to predict a contribution's quality, measured by the
success of build integration jobs.
We performed this analysis on a large corpus of open source projects
which incorporated continuous integration in their development practice. 
When a contribution to these projects passes integration tests, 
it suggests that the project developers
might have confidence in the changes. 
Thus, we use the commit's build integration status as a proxy for its quality.

@@

We present the approach we took to design our study by
following the Goal-Question-Metric (GQM) methodology~\cite{Basili84}:

@@

\begin{quotation}
\noindent\emph{Do social attributes
improve prediction of a contribution's build success when compared to
predictions based only on technical attributes?}
\end{quotation}

@@

\vspace{1ex}
\noindent\textbf{Metric.} We measured software contribution quality by the
build status of the
commit. We used a decision tree classifier~\cite{Quinlan86} to predict build 
success. Decision trees 
use binary decisions on the
attributes to predict a final class. They are well
established and interpretable. We used a Weka J48 decision tree for readability.
We first built a
decision tree classifier using only technical factors to 
predict build integration
success.  We then included the social factors, to 
identify the degree to which social factors improve the model's predictive
power.
%We evaluated the models with 10-fold cross-validation.

@@

The first five factors describe the changes in the commit.  The next two
factors cover the size of the project.  These factors proxy both the commit's complexity,
and the complexity of the project incorporating it.

\vspace{1ex}
\noindent\textbf{Social Factors.} We created a web scraper to extract social data
from developers' profiles.
We used the commit hash and the project name in the 
MSR Challenge~\cite{msr17challenge} data set to locate the
commit web page, leveraging \textit{github.com}'s URL standard for projects and
commit hashes.
%  using the syntax
% below. 
% \newline \newline
% \textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
% \newline \newline
%
We scraped the commit author's
\textit{username} from each commit page to
collect the following from the user's public profile:
%(\textit{https://github.com/\textbf{UserName}}). 
%

@@

\vspace{1ex}
\noindent\textbf{Compiling the data sets.} We compiled the gathered data into 
two data sets: one with only the technical
factors and the build status, and another with both technical and social
factors and the build status. We performed this split to analyze how well 
technical factors alone,
and then both technical and social
factors together predict build integration status.

@@

74.43\% of the builds in the reduced data set were successful, and thus a naive
classifier that always guesses ``successful'' would succeed 74.43\% of the time.
To penalize misclassification for
both classes equally, we created a second \emph{balanced} data set by randomly 
sub-sampling the successful builds until the two build status were
equal, through a process known as downsampling. We downsampled the entire
data set and then trained decision tree classifiers on both.  We validated the
classifiers with cross-validation. Since the cross-validation is stratified, the
50-50 proportion is maintained through the training and testing folds for the
downsampled data. We present results for both
data sets in Section~\ref{sec:successClassification},  and only the full data 
set in Sections~\ref{sec:core} and \ref{sec:followerSuccess}.

@@

\textbf{Kappa statistic}& 0.21 &  0.42 & 0.28  & 0.52 \\
\textbf{Average Precision}& 0.75 & 0.80  & 0.65  & 0.76 \\
\textbf{Average Recall}& 0.77 & 0.81 & 0.64 & 0.76 \\
\textbf{Average F-Measure}& 0.72 & 0.79  & 0.63  & 0.76 \\

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm.  Without downsampling, the 
decision tree shows an increase of 4.09\% \textit{correctly classified} instances when adding social data. On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, which outperformed the na\"ive
baseline model that 
guesses all builds passed (50.00\%).
After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement over the model that uses only the technical data.  These results show that
while technical data is helpful in predicting build integration success, social
data can be used to measurably improve predictive accuracy.

The precision, recall, and
F-measure are calculated over the successful and unsuccessful classes and then
averaged (standard in Weka).  All these measures, along with the Kappa
statistic, increase with social features, which provide evidence that social 
factors improve build success predictions. The mean 
absolute error also decreases, further substantiating the result.

@@

The decision tree results suggested that core team membership is
important to classifying build success (results not shown), and thus we investigated the question
directly. 
We classified the build success rate for
core team members and non-core team members using the data set without 
downsampling.  We found that core team members
triggered 89.39\% (331827/371208) of the builds in the data set; 75.58\%
(250808/331827) of those builds
passed.  Non-core team members triggered 10.61\% (39381/371208) of the builds 
in the data set,
and 64.70\% (25479/39381) of the builds passed.  The difference in success rate
is statistically significant (the p-value from a Fisher exact
test is 0.0000), suggesting that it is extremely unlikely to have occurred by
chance. Thus, in our dataset, core team 
members are more
likely to have a successful build integration than non-core team developers, likely
because core team members are naturally more familiar with the 
codebase.

@@

We also investigated whether developers with followers were more likely to have a
successful build integration, since followers could be seen as ``vouching'' for
a developer's abilities.  We therefore investigated the relationship between
follower count and build success rate (on the data without 
downsampling).  We first investigated whether developers with one or more followers
were more likely to pass the build than those without any followers at all. We
removed commits without any author profile
information (20374 builds, reducing the total build count to
350834). Only 0.87\% (3059/350834) of the
contributions in the data set were performed by developers with zero 
followers.  %, leaving 99.13\% (347775/350834) of the contributions 
%performed by developers with at least one follower.
Developers with 0 followers had a successful build commit 68.49\% (2095/3059)
of the time.  Developers with one or more followers were successful 74.78\%
(260056/347775) of the time.  The p-value of this difference, calculated with a
Fisher exact test, is 0.000, suggesting the difference in commit success rate between
developers with no followers and those with at least one follower is
very unlikely to have occurred by chance.
%
By contrast, developers with a large number of followers (1000 or more),
performed 6.94\% (24347/350834) of the contributions in the 
data set; their
contributions had a 78.61\% (19138/24347) success rate. The p-value of the
difference between this success rate and that of developers with no
followers is again 0.0000.  % The p-value means that that the difference between
% developers with 1000 followers and developers with no followers is very unlikely
% to occur by chance.

@@

%\paragraph{Internal validity}
In the study, we use Weka, a well known machine-learning toolkit,
to analyze our data.
Like all tools, Weka may contain bugs. 
Nonetheless, it is a community tested open source tool, which mitigates threats to 
internal validity. We also release our scripts, data, and decision tree 
results in the study.\footnote{https://github.com/mausotog/MSRChallenge2017/}

@@

The most similar related work to this paper is a case study investigating how
social and technical factors affect build
success on a project with 40 developers~\cite{Hassan06}.
We focus on a wide range of open source projects and
measure experience through developer profile information.
Other work has investigated the effects of various factors, such as team
organization, on predicting build success in a single
project~\cite{Kwan11,Wolf09}.
Previous investigations have established relationships between both social and 
technical 
attributes and 
different measures of success, such as pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse} or evaluation latency~\cite{Yu15}.
Tsay et al.~\cite{tsay14icse} 
show the importance of the social connection between the developer initiating a
pull request and the project integrators. We look at a broader aspect of 
social factors, and focus on predicting the Travis CI build success.
Gousios et al.~\cite{gousios14} found that the commit change location is
an important indicator to commit acceptance rate, and that including
test cases increases a
commit's acceptance rate~\cite{gousios15}.
Bettenburgh et 
al.~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, while Vasilescu et al.~\cite{vasilescu14} explore how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated into the automatic build process. Yu et al.~\cite{yu16} 
analyze the impact of CI usage in GitHub and its correlation with software 
quality. By contrast with their work, we focus on the relationship between social 
attributes and continuous integration success.

\balance
\section{Conclusion}

@@

% use section* for acknowledgement
\section*{Acknowledgments}
This material is based upon work supported by the National Science Foundation
Graduate Research Fellowship under Grant No. DGE-1252522.
The authors also gratefully acknowledge the partial support
of the AFRL (AFRL Contract No. FA8750-15-2-0075). Any 
findings or recommendations are those of the author(s) and do not
necessarily reflect the views of the sponsoring agencies.

@@

\bibliographystyle{plain}
\bibliography{MSRChallenge2017}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
